---
title: "ds.client.connection.server Packages users documentation"
author: "Patricia Ryser-Welch on Behalf of the DataSHIELD Team"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ds.client.connection.server Packages users documentation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# ds.client.connection.server
The purpose of this package is support DataSHIELD analysts and programmers to connect to some Opal servers in more supportive manner. The package offers some functions that can simplify certain operations within DataSHIELD. The latter includes:

1. Building some login data frames
2. Loging in and loging out of some DataSHIELD servers
3. Assigning values to some server variables
4. Executing some aggregates server-side functions 
5. Deleting some server-side variables
6. Checking some server-side variables

This package has been written to be used along the side the [dsBaseClient package](https://github.com/datashield/dsBaseClient).

# Installed packages
During the installation, this package will import the following packages: 

* DSI
* DSOpal
* RCurl
* httr
* R6
* opalr

These packages are required to connect to an Opal Server, so that some server-side variables can be manipulated and server-side functions can be executed. 



## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
